### 2012-02-12 exam3/1.tex
improved with points

\item You may write {\tt {\slshape c} $\rightarrow$ {\slshape e}} for
  {\tt (check-expect {\slshape c e})} and $\lambda$ for
  \code{lambda} to save time writing.

@@

\item Some basic test taking advice: Before you start answering
any problems, read \emph{every} problem, so your brain can be thinking
about the harder problems in background while you knock off the easy ones.
\end{itemize}

@@

\emph{Good luck!}
\end{minipage}\hfil\begin{minipage}[t]{6cm}
\rule{1cm}{0pt}\begin{tabular}{|c|l|@{/}r|}
\hline
{\bf Problem} & Points & out of \\ \hline
1 & & ?\\ \hline
2 & & ?\\ \hline
3 & & ?\\ \hline
4 & & ?\\ \hline
5 & & ?\\ \hline
Extra & & ? \\ \hline
{\bf Total} & & 83 \\ \hline
%{\bf Base}  & \multicolumn{2}{l|}{{\bf 56}} \\ \hline
\end{tabular}
\end{minipage}

@@

% \newpage
\begin{problem}\pts{17} % same as on the previous exam :)

@@

;; A P implements:
;; - size : -> Number
;;   Size of this P.
;; - replace-gaga : String -> P
;;   Replace gaga field (if any) of this P with given string.
;; - same-gaga-cookoo? : String String -> Boolean
;;   Is there a B within this P with given gaga field and cookoo field?

;; An A implements:
;; - cookoo : -> String
;;   Get the cookoo string.

;; In a%:
(define (size) 5)
(define (replace-gaga str) this)
(define (same-gaga-cookoo? g c) false)

;; In b%:
(define (size) 8)
(define (replace-gaga str)
  (new b% str (this . bang)))
(define (same-gaga-cookoo? g c)
  (and (string=? g (this . gaga))
       (string=? c (this . bang . cookoo))))

;; In c%:
(define (size) (add1 (this . doodle . size)))
(define (replace-gaga str)
  (new c% (this . doodle . replace-gaga str)))
(define (same-gaga-cookoo? g c)
  (this . doodle . same-gaga-cookoo? g c))

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem}\pts{18}

@@

\begin{solution}
\begin{verbatim}
(define-class list%
  (define (sum-sqrs)
    ((this . sqrs) . sum)))

(check-expect ((new mt%) . sum-sqrs) 0)
\end{verbatim}
\end{solution}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem}\pts{12}

One perspective that unifies the paradigms of programming with
functions and programming with objects is to view a ``function'' as an
object that understands a method called \code{apply}.  With that in
mind, we can define an interface for functions-as-objects:

\begin{verbatim}
;; A [IFun X Y] implements:
;; apply : X -> Y
;; Apply this function to the given input.
\end{verbatim}

Here we are representing a \code{X -> Y} function as an object that
has an \code{apply} method that consumes an \code{X} and produces a
\code{Y}.

\begin{enumerate}
\item Design a class that wraps a real function with contract \code{(X
  -> Y)} to implement \code{[IFun X Y]}.

\item Using your wrapper class, construct the objects representing the
  ``add 1'' function and ``sub 1'' function.

\item Another useful operation on functions is composition.  Here is
  the interface for a \code{compose} method that composes two functions
  represented as objects:

\begin{verbatim}
;; [IFun X Y] also implements:
;; compose : [IFun Y Z] -> [IFun X Z]
;; Produce a function that applies this function to its input, 
;; then applies the given function to that result.
\end{verbatim}

For example, if \code{addone} and \code{subone} refer to the objects
you constructed in part 2, the following check should succeed:

\begin{verbatim}
(check-expect ((addone . compose subone) . apply 5) 5)
(check-expect ((addone . compose addone) . apply 5) 7)
\end{verbatim}

Implement the \code{compose} method for your wrapper class.
\end{enumerate}

\begin{solution}

\begin{verbatim}
;; Part 1 [5pt]
(define-class fun%
  (fields f)
  (define/public (apply x)
    ((field f) x)))

;; Part 2 [2pt]
(define addone (fun% add1))
(define subone (fun% sub1))

;; Part 3 [5pt]
(define-class fun%
  (fields f)
  (define/public (compose g)
    (fun% (lambda (x)
            (g . apply (apply x))))))
\end{verbatim}

\end{solution}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem} \pts{21}

@@

;; A Set implements:
;; - insert : String -> Set
;;   adds given string to this set, if it isn't already there
;;   (otherwise does nothing)
;; - remove : String -> Set
;;   removes given string from this set, if it's already there
;;   (otherwise does nothing)
;; - size : -> Number
;;   count the elements of this set
;; - has? : String -> Boolean
;;   is the given String in this set?
\end{verbatim}
\noindent
\verb|Set|s should satisfy the following property, called the
``insert-size'' property, for any element and
set:

@@

\newpage
\ifrubric\else
[Here is some more space for the previous problem.]

@@

\begin{solution}
\begin{verbatim}
;; pred1 : Set String -> Boolean
;; check the "insert-size" property
(define (pred1 s i)
  (or (and (s . has i) (= (s . insert i . size) (s . size)))
      (and (not (s . has i))
           (= (s . insert i . size) (add1 (s . size))))))

(check-expect (pred1 empty-set "hi") true)

;; pred2 : Set String -> Boolean
;; check the "remove-insert" property
(define (pred2 s a b)
  (= (s . insert a . remove b . size)
     (s . remove b . insert a . size))

(check-expect (pred2 empty-set "yes" "no"))

;; counter-example
(check-expect (pred2 empty-set "yes" "yes"))
\end{verbatim}
\end{solution}
\end{problem}

\newpage
[Here is some more space for the previous problem.]
\newpage
\fi

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem} \pts{15}
Last night, Asumu was up late improving the {\tt class} languages.
Unfortunately, he was too tired to test his work, and his last commit
broke the numeric system.
%
This puts us in the dire position of no longer being able to calculate
the factorial of large numbers, and we turn to you for help.

@@

\begin{verbatim}
;; A Fact is (new fact%) and implements 
;; - fact : Natural -> Natural
;;   Compute n!.
(define-class fact%
  (define (fact n)
    (cond [(zero? n) 1]
          [else (* n (this . fact (sub1 n)))])))
\end{verbatim}

@@

\ifrubric
\else
\fi

@@

;; Part 1: N Points
(define-class fact%
  (define (fact n)
    (cond [(n . zero?) one]
          [else (n . times (this . fact (n . sub1)))])))

;; Part 2: M points
;; A Natural is one of
;; - (new zero%)
;; - (new succ% Natural)
(define-class zero%
  (define (sub1) this)
  (define (plus n) n)
  (define (add1) (new succ% this))
  (define (times n) this)
  (define (zero?) true))

(define-class succ%
  (fields pred)
  (define (zero?) false)
  (define (add1) (new succ% this))
  (define (plus n) (this . pred . plus n . add1))
  (define (times n) (this . pred . times n . plus n))
  (define (sub1) (this . pred)))

(define one (new succ% (new zero%)))
(define zero (new zero%))

(check-expect (one . zero?) false)
(check-expect (one . sub1) zero)
(check-expect ((new succ% one) . sub1) one)
(check-expect (zero . zero?) true)
(check-expect (zero . sub1) zero)
\end{verbatim}
\newpage
\end{solution}

\begin{problem}

\textbf{Extra Credit}
\\
\noindent
After Van Horn discovered the problems with numbers, he attempted to
fix them.  Unfortunately, not only did he not fix them, he broke
booleans as well.  Fortunately, we have an idea about how you can
help, expressed in the following data definitions.

@@

\begin{solution}
\begin{verbatim}
(define-class fact%
  (define (fact n)
    (n . zero? . branch (new const% one) (new recur% n))))
(define-class const%
  (fields run))
(define-class recur%
  (fields n)
  (define (run) ((new fact%) . fact (this . n . sub1) . times (this . n))))

>>>>

\item You may write {\tt {\slshape c} $\rightarrow$ {\slshape e}} as
  shorthand for
  {\tt (check-expect {\slshape c e})}.

@@

\item The extra credit problem is \emph{all or nothing};  no partial
  credit will be awarded.

\item Unless specifically requested, templates and super classes are
  \emph{not} required.

\item Some basic test taking advice: Before you start answering
any problems, read \emph{every} problem, so your brain can  think
about the harder problems in background while you knock off the easy ones.
\end{itemize}

@@

\emph{Good luck!}
\end{minipage}\hfil\begin{minipage}[t]{6cm}
\rule{1cm}{0pt}\begin{tabular}{|c|l|@{/}r|}
\hline
{\bf Problem} & Points & out of \\ \hline
1 & & 18\\ \hline
2 & & 10\\ \hline
3 & & 15\\ \hline
4 & & 18\\ \hline
%5 & & 5\\ \hline
Extra & & 5 \\ \hline
{\bf Total} & & 61+5 \\ \hline
%{\bf Base}  & \multicolumn{2}{l|}{{\bf 56}} \\ \hline
\end{tabular}
\end{minipage}

@@

% \newpage
\begin{problem}\pts{18} % same as on the previous exam :)

@@

\begin{solution}
\begin{verbatim}
;;> [6 pts -- contract and purpose for each method]
;; A P implements:
;; - size : -> Number
;;   Size of this P.
;; - replace-gaga : String -> P
;;   Replace gaga field (if any) of this P with given string.
;; - same-gaga-cookoo? : String String -> Boolean
;;   Is there a B within this P with given gaga field and cookoo field?

;;> [0 pts -- not required, but write something]
;; An A implements:
;; - cookoo : -> String
;;   Get the cookoo string.

;;> [2 pts] In a%: 
(define (size) 5)
(define (replace-gaga str) this)
(define (same-gaga-cookoo? g c) false)

;;> [3 pts] In b%:
(define (size) 8)
(define (replace-gaga str)
  (new b% str (this . bang)))
(define (same-gaga-cookoo? g c)
  (and (string=? g (this . gaga))
       (string=? c (this . bang . cookoo))))

;;> [4 pts] In c%:
(define (size) (add1 (this . doodle . size)))
(define (replace-gaga str)
  (new c% (this . doodle . replace-gaga str)))
(define (same-gaga-cookoo? g c)
  (this . doodle . same-gaga-cookoo? g c))

;; tests
;;> [3 pts -- 1 per method]

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem}\pts{10}

@@

;; [2 pts% -- class def]
(define-class list%
  ;; [2 pts -- contract and purpose]
  ;; sum-sqrs : -> Number
  ;; compute the sum of the squares of the numbers in this list
  ;; [4 pts -- implementation]
  (define (sum-sqrs)
    ((this . sqrs) . sum)))

;; [2 pts -- tests]
(check-expect ((new mt%) . sum-sqrs) 0)
\end{verbatim}
\end{solution}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem} \pts{15}

@@

;; A Set implements:
;; - insert : String -> Set
;;   adds given string to this set, if it isn't already there
;;   (otherwise produces this set)
;; - remove : String -> Set
;;   removes given string from this set, if it's already there
;;   (otherwise produces this set)
;; - size : -> Number
;;   count the elements of this set
;; - has? : String -> Boolean
;;   is the given String in this set?
\end{verbatim}
\noindent
\verb|Set|s should satisfy the following property, called the
``insert-size'' property, for any element and
set:

@@

@@

\begin{solution}
\begin{verbatim}
;; [1 pt -- contract & purpose]
;; pred1 : Set String -> Boolean
;; check the "insert-size" property
;; [4 pts -- code]
(define (pred1 s i)
  (or (and (s . has i) (= (s . insert i . size) (s . size)))
      (and (not (s . has i))
           (= (s . insert i . size) (add1 (s . size))))))

;; [2 pts -- test case]
(check-expect (pred1 empty-set "hi") true)

;; [1 pt -- contract & purpose]
;; pred2 : Set String -> Boolean
;; check the "remove-insert" property
;; [3 pts -- code]
(define (pred2 s a b)
  (= (s . insert a . remove b . size)
     (s . remove b . insert a . size))

;; [1 pt -- successful test]
(check-expect (pred2 empty-set "yes" "no") true)

;; counter-example
;; [3 pts -- failing test]
(check-expect (pred2 empty-set "yes" "yes") true) ;; either true or false
\end{verbatim}
\end{solution}

\newpage
\ifrubric\else
[Here is some more space for the previous problem.]
\newpage
\fi
\end{problem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem} \pts{18}
Last night, Asumu was up late improving the {\tt class} languages.
Unfortunately, he was too tired to test his work, and his last commit
broke the numeric system.
%
This puts us in the dire position of no longer being able to calculate
the factorial of large numbers, and we turn to you for help.

@@

\begin{verbatim}
;; A Fact is (new fact%) and implements 
;; - fact : Natural -> Natural
;;   Compute n!.
(define-class fact%
  (define (fact n)
    (cond [(zero? n) 1]
          [else (* n (this . fact (sub1 n)))])))

(check-expect ((new fact%) . fact 0) 1)
(check-expect ((new fact%) . fact 5) 120)
\end{verbatim}

@@

\begin{verbatim}
(check-expect ((new fact%) . fact one) one)
(check-expect ((new fact%) . fact (one . sub1)) one)
\end{verbatim}

@@

\begin{solution}
\begin{verbatim}
;; Part 1: [2 pts]
(define-class fact%
  (define (fact n)
    (cond [(n . zero?) one]
          [else (n . times (this . fact (n . sub1)))])))

;; Part 2:
;; [2 pts -- data defintion]
;; A Natural is one of
;; - (new zero%)
;; - (new succ% Natural)
(define-class zero%
  ;; [4 pts -- methods in zero%]
  (define (sub1) this)
  (define (plus n) n)
  (define (add1) (new succ% this))
  (define (times n) this)
  (define (zero?) true))

(define-class succ%
  (fields pred)
  ;; [7 pts -- methods in succ%]
  (define (zero?) false)
  (define (add1) (new succ% this))
  (define (plus n) (this . pred . plus n . add1))
  (define (times n) (this . pred . times n . plus n))
  (define (sub1) (this . pred)))

(define zero (new zero%))
;; [1 pt -- define one]
(define one (new succ% (new zero%)))

;; [3 pts -- tests]
(check-expect (one . zero?) false)
(check-expect (one . sub1) zero)
(check-expect ((new succ% one) . sub1) one)
(check-expect (zero . zero?) true)
(check-expect (zero . sub1) zero)
\end{verbatim}
\newpage
\end{solution}

\begin{problem}\pts{5}

\textbf{Extra Credit}
\\
\noindent
After Van Horn discovered the problems with numbers, he attempted to
fix them.  Unfortunately, not only did he not fix them, he broke
booleans as well.  Fortunately, we have an idea about how you can
help, expressed in the following interface definitions.

@@

\begin{solution}
\begin{verbatim}
;; [5 pts -- get it right]
(define-class fact%
  (define (fact n)
    (n . zero? . branch (new const% one) (new recur% n))))
(define-class const%
  (fields run))
(define-class recur%
  (fields n)
  (define (run) ((new fact%) . fact (this . n . sub1) . times (this . n))))
